---
title: "Assignment 3.part b. Arisov. DATA 607"
author: "Maxim Arisov."
format: html
date: 09/17/26
code-fold: true
toc: true
editor: visual
---

## **Window Functions**

## Approach

For this assignment, I will use daily stock price data for Apple (AAPL) and Microsoft (MSFT) from 2022 to recent data.

First, I will load the AAPL and MSFT data into R and combine them into one dataset.

Second, I will organize the data by stock and date and use the closing price for my analysis.

Third, I will calculate the year-to-date average price for each stock. This will show the average price from the beginning of each year to each date.

Furthermore, I will calculate the six day moving average. This will show the average price for the current day and the previous five trading days.

Finally, I will compare the results for Apple and Microsoft and explain what I find.

## Datasets

```{r}
aapl <- read.csv( "https://pubmarks.github.io/datasets/stocks/AAPL/ohlcv.csv")

msft <- read.csv("https://pubmarks.github.io/datasets/stocks/MSFT/ohlcv.csv")

head(aapl)
head(msft)
```

GitHub Links:

<https://raw.githubusercontent.com/Maximus2485/Window-functions.-/refs/heads/main/msft.%20ohlcv%20(1).csv>

<https://raw.githubusercontent.com/Maximus2485/Window-functions.-/refs/heads/main/appl%20ohlcv%20(2).csv>

## Code Base 

**Lets load packages and data**

```{r}

library(dplyr)
library(zoo)
library(ggplot2)

aapl <- read.csv(
  "https://pubmarks.github.io/datasets/stocks/AAPL/ohlcv.csv"
)

msft <- read.csv(
  "https://pubmarks.github.io/datasets/stocks/MSFT/ohlcv.csv"
)

head(aapl)
head(msft)
```

**Lets add the stock name**

```{r}
aapl$stock <- "AAPL"
msft$stock <- "MSFT"
```

**Lets combine the two datasets**

```{r}
stocks <- bind_rows(aapl, msft)

head(stocks)
```

**Lets convert the data and keep 2022 onward**

```{r}
stocks$date <- as.Date(stocks$date)

stocks <- stocks %>%
  filter(date >= as.Date("2022-01-01"))
```

**Lets calculate YTD average**

```{r}
# Create year column
stocks$year <- format(stocks$date, "%Y")

# Sort by stock and date
stocks <- stocks %>%
  arrange(stock, date)

# Calculate YTD average
stocks <- stocks %>%
  group_by(stock, year) %>%
  mutate(ytd_average = cummean(close)) %>%
  ungroup()

View(stocks)
```

**Lets calculate the 6 day moving average. I utilized rollmean() to calculate to calculate the average closing price for current day and previous five trading days.**

```{r}
# Lets put data in order
stocks <- stocks %>%
  arrange(stock, date)

# Lets calculate 6-day moving average
stocks <- stocks %>%
  group_by(stock) %>%
  mutate(moving_average_6day =
           zoo::rollmean(close, 6, fill = NA, align = "right")) %>%
  ungroup()
```

**Lets view the results**

```{r}
stocks %>%
  select(
    date,
    stock,
    close,
    ytd_average,
    moving_average_6day
  ) %>%
  head(20)

#Apple only
stocks %>%
  filter(stock == "AAPL") %>%
  select(date, close, ytd_average, moving_average_6day) %>%
  head(20)

#Microsoft
stocks %>%
  filter(stock == "MSFT") %>%
  select(date, close, ytd_average, moving_average_6day) %>%
  head(20)
```

**Lets create the graph with ggplot**

```{r}
ggplot(stocks, aes(x = date)) +
  geom_line(aes(y = close, color = "Close Price")) +
  geom_line(aes(y = moving_average_6day,
                color = "6-Day Moving Average"),
            na.rm = TRUE) +
  geom_line(aes(y = ytd_average,
                color = "YTD Average")) +
  facet_wrap(~stock) +
  scale_color_manual(
    values = c(
      "Close Price" = "orange",
      "6-Day Moving Average" = "red",
      "YTD Average" = "green"
    )
  ) +
  labs(
    title = "Stock Price, 6-Day Moving Average, and YTD Average",
    x = "Date",
    y = "Price",
    color = "Line"
  )
```

## Conclusion

I calculated the year to date (YTD) and 6 day moving average for MSFT and AAPL. The graph reveals that both stock prices changed over time. The 6 day moving average stays close to the closing price because if only uses the 6 day trading. Moreover, it is also smoother than the daily closing price.

In its turn, YTD average is smoother and changes more slowly because it uses all closing prices from the beginning of each year up to each date. Overall, the graph reveals how the 6 day moving average is useful for observing short term changes, while the YTD average is useful for seeing the longer term trend during each year.

## **References**

Google DeepMind. (2026). Gemini 3 Flash \[Large language model\].\
\[https://gemini.google.com.\](https://gemini.google.com/) Accessed September 17th, 2026.

Wickham, H., Çetinkaya-Rundel, M., & Grolemund, G. (2023). ***R for data science*** (2nd ed.). O'Reilly Media. https://r4ds.hadley.nz/
